
# Introduction to WDL


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

There is ongoing work on WDL the specification, thus it has multiple versions.  Currently there are three versions to note:
- draft-2 - this version was the version that much of the Broad's public facing documentation and example workflows were written in. 
- 1.0 - this is a more recent version that is the most up to date version of the specification that Cromwell can interpret.  We use WDL 1.0 at the Hutch.
- 1.1 - this is an even more recent version that not all WDL engines support yet. 

We'll be using WDL 1.0 in this course but you can always check out the [openwdl repo](https://github.com/openwdl/wdl) for more information about tweaking these instructions for different versions of WDL.  

## WDL 1.0 Syntax

```{r, include = TRUE}
cow::borrow_chapter(repo_name = "openwdl/wdl-docs", "Docs/Base_structure.md", branch = "1.0")
```

## Tools for Editing WDLs
- VSCode has an extension
- ways to plot a DAG? 


## Workflow Structure


### Inputs


### Tasks


#### Inputs

#### Command

#### Outputs

#### Runtime


### Managing Software Environments


## Wiring Tasks Together

